Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-09-11 writedown/period.noneq.msm/paper.tex
further revise

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{red} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
% \newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\vect}[1]{#1}
\newcommand{\dd}[1]{\textsf{#1}}
\newcommand{\fwd}[0]{\textrm{fw}}
\newcommand{\bwd}[0]{\textrm{bw}}
\newcommand{\period}[0]{T_{\textrm{P}}}
\newcommand{\ml}[0]{\mathcal {L}}
\newcommand{\mo}[0]{\mathcal {O}}
\newcommand{\mbp}[0]{\mathbb {P}}
\newcommand{\mc}[0]{\mathcal {C}}
\newcommand{\dt}[0]{\Delta t}
\newcommand{\id}{\mathrm{Id}}
% \newcommand{\myphi}{\boldsymbol\Phi}
% \newcommand{\mymu}{\boldsymbol\mu}
\newcommand{\myphi}{\Phi}
\newcommand{\mymu}{\mu}
\newcommand{\prob}{\textrm{P}}

@@

\title{Building Markov State Model for a Periodically Driven Non-Equilibrium System}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Zuse Institut Berlin, Germany}
\author{Christof Sch\"utte}
\email{Christof.Schuette@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institut Berlin, Germany}

@@

Current achievement of MSM in equilibrium cases.

@@

\section{Discretization of the non-equilibrium molecular dynamics}
\label{sec:disc}

We consider  the following SDE form of the externally driven MD:
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
where the configurational space is denoted by $\Omega$, and the variable by
$\vect x$.   $V(\vect x)$ is the molecular interaction potential and
$E(t)D(x_t)$ the time-dependent external perturbation with the
$T$-periodic external field $E(t)$. $\beta$ is the inverse temperature,
i.e.~$\beta = 1/(k_B\mathcal T)$.
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$ is governed by
Fokker-Planck equation:
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
where $\ml^\dagger(t)$ is the adjoint of the generator
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator wrt to $\vect x$. 
The periodicity of the external driven indicates the periodicity of the generator,
i.e.~$\ml(t) = \ml(t+T)$.  Now
introduce a partition of the configurational space $\Omega$ into finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_n\}$, which satisfy $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Following Ref.~\cite{latorre2011structure}, the Fokker-Planck Eq.~\ref{eq:disc-fp}
is discretized, which results in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, n\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{n\times n}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^n L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
for all real time $t\geq 0$.
{Moreover,
the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.}
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability densities according to the associated Master equation
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^T(t)\cdot \vect p(t)
where $\vect L(t)^T$ denotes the matrix transpose of $\vect L(t)$ and
$p_i(t)$, for example, the probability to be in state $i$ at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p_i(0)\ge 0$ componentwise, then $p_i(t)\ge 0$ and $\sum_i
p_i(t) = \sum_ip_i(0)$.
% The solution of the master equation need no be
% periodic.
It can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{n\times n}$ that solves
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^T(t)\myphi(t), \quad \myphi(0) = \id
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^n \Phi_{ij}(t)  =  1$,
that is, $\myphi^T(t)$ is a
stochastic matrix satisfying $\myphi^T(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^T\in \mathbb R^n$.
Regarding these considerations, we find
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.
 The
discretization sets that we used to go from $\vect x_t$ and $\ml(t)$ to $\vect X_t$
and $\vect L(t)$, respectively, can be assumed to provide an arbitrarily fine
partition of the original state space; then the transport properties
of $\vect L(t)$ are almost perfect approximations of the transport properties
of $\ml(t)$, in particular $p_i(t)=\prob(x_t\in \Omega_i)$.

\section{Floquet theory}
\label{sec:floquet}

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. $\vect Y$ satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L(t+T)^T \vect Y(t)=\vect L(t)^T\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T]$.  
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0).
\end{align}

Since $\myphi(T)$ is a stochastic matrix, the spectrum $\sigma(\myphi(T))$
of the matrix $\myphi(T)$ is contained in the circle in
the complex plane, i.e., each eigenvalue $\lambda\in \sigma(\myphi(T))$
satisfies $|\lambda|\le 1$. Furthermore $1\in\sigma(\myphi(T))$ is an
eigenvalue with left eigenvector $\vect e$ and a right eigenvector $\mymu$
satisfying
$\myphi(T)\mymu=\mymu$.
From now on, we assume $\myphi(T)$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
\begin{align}
\label{eq:floq-14}  
\myphi(mT) \mymu = \mymu,\quad m=0,1,2,\ldots,
\end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
\begin{align}
\label{eq:floq-15}  
\myphi(mT)\vect p(0)\to \mymu,\quad m\to\infty.
\end{align}

Using the Floquet theorem, the time-inhomogeneous Markov process $X_t$
is simplified into a \emph{time-homogeneous} (not necessarily
reversible) Markov jump process $\tilde X_{m} = X_{mT}, \ m\in\mathbb
N$.
Therefore, comparing with $X_t$, we prefer to consider the process $\tilde X_{m}$
that helps understand the original dynamics $x_t$, because the powerful theories and
computational tools for time-homogeneous processes can be directly applied.
There is no doubt that information within one period is lost by using this temporal
discretization, however, information regarding the long-term behaviors
of the system would be satisfactorily described as long as their time-scales
are much longer than the period.
At the same time,
the computational cost of generating $\tilde X_{m}$ is mush less demanding
than the brute force simulations of NEMD, which implies lower
statistically uncertainty in calculating the observables of interest.
% due to the discretization in both the spacial and temperal directions.
% In the numercal example in Sec.~\ref{sec:alanine}, we
% compute the

\section{Building the Markov state model}
\label{sec:build-msm}

If the discretization cells $\Omega_i$, $i=1,\ldots,n$ form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix $\vect P=\myphi^T(T)$ still is a fine-scale description of the
dynamics.  Now we want to coarse our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $k\ll n$
macrostates: The resulting $k\times k$ MSM transition matrix $\hat{\vect P}$
then defines the coarse grained long term kinetics that should
approximate the original long term kinetics well. In Ref.~\cite{sarich2010approximation}
and \cite{schuette2011markov} it
has been shown how to do this if $\vect P$ satisfies the detailed balance
condition: (1) Identify the cores of the metastable sets of the
dynamics, (2) use them as milestones to construct an MSM in which the
macrostates are the metastable core sets and $\hat{\vect P}$ is the transition
matrix of the milestone process that models the jumping behavior of
the original dynamics between the metastable regions.

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach recently proposed in Ref.~\cite{sarich2014utilizing}.
It allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $k$ core sets $C_1,\ldots, C_k\subset S$, and we denote $C=S\setminus\cup_j C_j$. Following
Ref.~\cite{djurdjevac2010markov}, Thm. 3.1, the coarse grained transition
matrix $\hat{P}$ has to be computed as follows:
\begin{enumerate}
\item For the process associated with $P$ compute the forward and backward committors $q^\fwd_j$ and $q^\bwd_j$  for each core set $C_j$. This can be done by solving the linear equations
\begin{align}
(P-\id) q^\fwd_j(i) & =  0, \quad i\in C\\
q^\fwd_j(i) & =  1,\quad i\in C_j\\
q^\fwd_j(i) & =  0,\quad i\in C_l,l\not=j
\end{align}
and
\begin{align}
(P^b-\id) q^\bwd_j(i) & =  0, \quad i\in C\\
q^\bwd_j(i) & =  1,\quad i\in C_j\\
q^\bwd_j(i) & =  0,\quad i\in C_l,l\not=j
\end{align}
where $P^b$ denotes the transition matrix of the time-reversed process
given by $P^b_{ji}=\mu_i P_{ij}/\mu_j$.

Alternatively, the forward and
backward committors can be sampled via the definitions: The forward
committor $q^\fwd_j(i)$ is defined as the probability of visiting coreset
$C_j$ next conditioned on being at state $i$.  The
backward committor $q^\bwd_j(i)$  is defined as the probability
of last coming from $C_j$ conditioned on being at state $i$.
\item Compute $\hat{\mu}(j)=\sum_{i\in S} q^\bwd_j(i)\mu(i)$,
  which is the invariant measure of the MSM
\item Construct the MSM transition matrix $\hat{P}$ according to
  \begin{align}
    \label{eq:msm-tmatrix}
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u_i v_i \mu_i$.
\end{enumerate}

@@

>>>>

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{red} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
% \newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\vect}[1]{#1}
\newcommand{\dd}[1]{\textsf{#1}}
% \newcommand{\fwd}[0]{\textrm{fw}}
% \newcommand{\bwd}[0]{\textrm{bw}}
\newcommand{\fwd}[0]{+}
\newcommand{\bwd}[0]{-}
\newcommand{\period}[0]{T_{\textrm{P}}}
\newcommand{\ml}[0]{\mathcal {L}}
\newcommand{\mo}[0]{\mathcal {O}}
\newcommand{\mbp}[0]{\mathbb {P}}
\newcommand{\mc}[0]{\mathcal {C}}
\newcommand{\dt}[0]{\Delta t}
\newcommand{\id}{\mathrm{Id}}
% \newcommand{\myphi}{\boldsymbol\Phi}
% \newcommand{\mymu}{\boldsymbol\mu}
\newcommand{\myphi}{\Phi}
\newcommand{\mymu}{\mu}
\newcommand{\prob}{\textrm{P}}

@@

\title{Building Markov State Models for Periodically Driven Non-Equilibrium Systems}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Zuse Institut Berlin, Germany}
\author{Christof Sch\"utte}
\email{Christof.Schuette@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institut Berlin, Germany}

@@

Current achievement of MSM in equilibrium cases:
\begin{itemize}
\item well developed in theory and applications \cite{A19-31, A19-1}
\item software for MSM building \cite{A19-49, MSMBuilder}
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Non-equilibrium molecular dynamics and its discretization}
\label{sec:disc}
For the sake of simplicity we consider diffusive molecular dynamics in an energy landscape $V$ driven by the time-dependet external driving force $E(t)D(x_t)$ with the
$T$-periodic external field $E(t)$:
\begin{align}
  \label{eq:disc-1}
  d\vect x_t = \Big(-\nabla V(\vect x_t) + E(t) D(\vect x_t)\Big)dt + \sqrt{2\beta^{-1}} d\vect w_t, 
\end{align}
where $x_t\in\Omega$ denotes the state of the molecular system at time $t$ in state space $\Omega$, $\vect w_t$ denotes standard $n$-dimensional Brownian motion, and $\beta$ the inverse temperature,
i.e.~$\beta = 1/(k_B\mathcal T)$.
The propagation of probability
densities $\rho=\rho(\vect x,t)$ based on this kind of dynamics in the sense
of $\rho(\vect x,t)dx=\prob[\vect x_t\in [\vect x,\vect x+d\vect x)]$ is governed by
Fokker-Planck equation:
\begin{align}
  \label{eq:disc-fp}
  \frac{\partial \rho}{\partial t} = \ml^\dagger(t) \rho,
\end{align}
where $\ml^\dagger(t)$ is the adjoint of the generator
\begin{align}
  \label{eq:disc-3}
  \ml(t)=\beta^{-1}\Delta_{\vect x}+\Big(-\nabla_x V(\vect x) + E(t)D(\vect x)\Big)\cdot\nabla_{\vect x},
\end{align}
where $\Delta_{\vect x}$ denotes the Laplacian operator and $\nabla_{\vect x}$
the nabla-operator wrt to $\vect x$. 
The periodicity of the external driving force induces the periodicity of the generator,
i.e.~$\ml(t) = \ml(t+T)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Spatial discretization: Master equation}
We will now introduce an appropriate spatial discretization of this kind of non-equilibirium MD -- this is done for reasons of simplicity only; we could completely avoid it for the price of more technical arguments.
For achieving this discretization, we 
introduce a partition of state space $\Omega$ into a finite number of disjoint
sets $\{ \Omega_1, \cdots, \Omega_n\}$ satisfying $\Omega = \cup_i \Omega_i$,
$\Omega_j\cap \Omega_j = \emptyset,\ \forall i\neq j$.
Utilizing the procedure described in Ref.~\cite{latorre2011structure} the original Fokker-Planck equation~(\ref{eq:disc-fp})
is discretized, resulting in a time-inhomogeneous Markov jump process in state
space $S = \{1, \cdots, n\}$ with time-dependent rate
matrix $\vect L(t) \in \mathbb R^{n\times n}$ satisfying
\begin{align}\label{eq:disc-4}
\sum\limits_{j=1}^n L_{ij}(t) & =  0\\ \label{eq:disc-5}
L_{ij}(t) & \ge  0, \quad i\not= j\\
L_{ij}(t) & =  L_{ij}(t+T)
\end{align}
for all real time $t\geq 0$.
Moreover, the rate matrix $L$ has the form $\vect L(t)=\vect L_0+E(t)\vect L_1$
where $E(t)$ is periodic with period $T>0$.
In analogy to \eqref{eq:disc-fp}, the Markov jump process generated by
$\vect L(t)$ transports probability distributions according to the associated Master equation
\begin{align}
  \label{eq:disc-master}
  \frac{d\vect p(t)}{dt} = \vect L^T(t)\cdot \vect p(t)
\end{align}
where $\vect L(t)^T$ denotes the matrix transpose of $\vect L(t)$, $\vect p(t)$ is an $n$-vector denoting the probability distribution on $S$ at time $t$, and
$p_i(t)$, for example, the probability to be in state $i$ (which corresponds to set $\Omega_i$) at time $t$.
As usual the properties (\ref{eq:disc-4}) and (\ref{eq:disc-5}) of
$\vect L(t)$ guarantee that the total probability mass is conserved,
i.e., if $p_i(0)\ge 0$ componentwise, then $p_i(t)\ge 0$ and $\sum_i
p_i(t) = \sum_ip_i(0)$.
% The solution of the master equation need no be
% periodic.
The temporal evolution of the probability distribution $\vect p(t)$ can be formally written
\begin{align}  \label{eq:disc-8}
\vect p(t)=\myphi(t)\vect p(0)
\end{align}
by using the
associated propagator matrix $\myphi(t)\in\mathbb R^{n\times n}$ that solves
\begin{align}
  \label{eq:disc-master-phi}
  \frac{d}{dt}\myphi(t) = \vect L^T(t)\myphi(t), \quad \myphi(0) = \id .
\end{align}
Since the last equation can be considered column-wise, the propagator matrix inherits column-wise conservation properties:
$\Phi_{ij}(t) \ge  0$
and $\sum\limits_{i=1}^n \Phi_{ij}(t)  =  1$,
that is, $\myphi^T(t)$ is a
stochastic matrix satisfying $\myphi^T(t)\vect e=\vect e$
with $\vect e=(1,\ldots,1)^T\in \mathbb R^n$.
Regarding these considerations, we find
\begin{align}
\label{eq:disc-10}  
\myphi_{ij}(t)=\prob\left(\vect X_t=i\mid \vect X_0=j \right),
\end{align}
where $\vect X_t$ denotes the Markov process generated by $\vect L(t)$.

The discretization sets that we used to go from $\vect x_t$ and $\ml(t)$ to $\vect X_t$
and $\vect L(t)$, respectively, can be assumed to provide an arbitrarily fine
partition of the original state space; then the transport properties
of $\vect L(t)$ are almost perfect approximations of the transport properties
of $\ml(t)$, in particular the approximation $p_i(t)\approx \prob(x_t\in \Omega_i)$ is almost perfect.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Temporal discretization: Floquet theory}
\label{sec:floquet}

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. It satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L(t+T)^T \vect Y(t)=\vect L(t)^T\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T)$.  
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0),
\end{align}
where $\myphi^m(T)$ denotes the $m$th power of the matrix $\myphi(T)$.
Since $\myphi(T)$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with left eigenvector $\vect e=(1,\ldots,1)^T$ and a right eigenvector $\mymu$
satisfying
$\myphi(T)\mymu=\mymu$.
From now on, we assume $\myphi(T)$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
\begin{align}
\label{eq:floq-14}  
\myphi(mT) \mymu = \mymu,\quad m=0,1,2,\ldots,
\end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
\begin{align}
\label{eq:floq-15}  
\myphi(mT)\vect p(0)\to \mymu,\quad m\to\infty,
\end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.

Using the Floquet theorem, the time-inhomogeneous Markov jump process $X_t$
is simplified into a \emph{time-homogeneous} (not necessarily
reversible) Markov jump process $\tilde X_{m} = X_{mT}, \ m\in\mathbb
N$.
Therefore, we prefer to consider the discrete-time process $\tilde X_{m}$ instead of the time-continuous process $X_t$
because the powerful theories and
computational tools for time-homogeneous Markov processes can be directly applied. Many of these tools require the transition matrix $\myphi(T)$ to satisfy the detailed balance condition. The computations in the Appendix show that $\myphi$ will in general \emph{not} satisfy this condition; in fact  the deviation form reversibility can be estimated from the work of the periodic driving does to the system.

There is no doubt that information within one period is lost by using this temporal
discretization, however, information regarding the long-term behavior of the system on time-scales
much longer than the period will be perfectly described because of $\tilde X_{m} = X_{mT}\approx x_{mT}$ whenever our spatial discretization is fine enough.
At the same time,
the computational cost of generating $\tilde X_{m}$ is much less demanding
than the brute force simulations of NEMD, which implies lower
statistically uncertainty in calculating the observables of interest.

% due to the discretization in both the spacial and temperal directions.
% In the numercal example in Sec.~\ref{sec:alanine}, we
% compute the

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Markov State Model}
\label{sec:build-msm}

% \subsection{Building the Markov State Model from the Floquet transition matrix}
If the discretization cells $\Omega_i$, $i=1,\ldots,n$ form a fine
partition of the molecular state space, the Markov chain defined via
the transition matrix 
\begin{equation}\label{P}
\vect P=\myphi^T(T)
\end{equation}
 is discrete in time but in space it still is a fine-scale description of the transport properties of the
dynamics with a very large number $n$ of states.  Now we want to coarsen our description much further by
constructing a Markov State Model (MSM) for $\vect P$ with $k\ll n$
\emph{macrostates} that should be the metastable states of the system: The resulting $k\times k$ MSM transition matrix $\hat{\vect P}$
then defines the coarse grained long term kinetics that shall
approximate the original long term kinetics well. 
The idea behind MSM building is that given the molecular system under consideration exhibits metastable conformations then it is usually possible to construct a
relatively small number of discrete sets --the metastable sets that form the so-called macrostates-- that
correctly describe the slow  dynamics, and 
in each set the fast dynamics relaxes on some timescale significantly shorter than the metastable timescales.
Then if the MSM dynamics reproduces the slow timescales and the corresponding transitions
of the original dynamics~\eqref{eq:disc-1},
the former is considered to be a good approximation of the later.

MSM building has been attracted a lot of attention recently, and theory  \cite{A19-31} as well as algorithms \cite{A19-1}, applications (see e.g. \cite{A19-26,PNAS09} for two of hundreds of articles) and software \cite{A19-49, MSMBuilder} have been developed to quite an extend. However, 
by far most of the literature is related to building \emph{standard} MSMs for equilibirium MD. In standard MSM also the transition region has to be discretized, a feature that often forces the user to incorporate more macrostates than essentially needed to approximate the long-term kinetics.
In Ref.~\cite{sarich2010approximation, A19-31,schuette2011markov,BucheteHummer} it has been shown how to construct \emph{non-standard} MSM that avoid this problem for equilibrium MD, i.e., if $\vect P$ satisfies the detailed balance
condition: (1) Identify the cores of the metastable sets of the
dynamics, (2) use them as milestones to construct an MSM in which the
macrostates are the metastable core sets and $\hat{\vect P}$ is the transition
matrix of the milestone process \cite{A19-31,schuette2011markov,A19-29} that models the jumping behavior of
the original dynamics between the metastable regions.

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing}.
It allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $k$ core sets $C_1,\ldots, C_k\subset S$, and we denote $C=S\setminus\cup_j C_j$. Following
Ref.~\cite{djurdjevac2010markov}, Thm. 3.1, the coarse grained transition
matrix $\hat{P}$ then has to be computed as follows:
\begin{enumerate}
\item For the process associated with $P$ compute the forward and backward committors $q^\fwd_j$ and $q^\bwd_j$  for each core set $C_j$. This can be done by solving the linear equations
\begin{align}
(P-\id) q^\fwd_j(i) & =  0, \quad i\in C\label{qfwd}\\
q^\fwd_j(i) & =  1,\quad i\in C_j\nonumber\\
q^\fwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
and
\begin{align}
(P^b-\id) q^\bwd_j(i) & =  0, \quad i\in C\label{qbwd}\\
q^\bwd_j(i) & =  1,\quad i\in C_j\nonumber \\
q^\bwd_j(i) & =  0,\quad i\in C_l,l\not=j\nonumber
\end{align}
where $P^b$ denotes the transition matrix of the time-reversed process
given by $P^b_{ji}=\mu_i P_{ij}/\mu_j$.
\item Compute $\hat{\mu}(j)=\sum_{i\in S} q^\bwd_j(i)\mu(i)$,
  which is the invariant measure of the MSM
\item Construct the MSM transition matrix $\hat{P}$ according to
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u_i v_i \mu_i$.
\end{enumerate}

\noindent\textbf{Remark 1:}
Alternatively, the forward and
backward committors can be \emph{sampled} by means of the following procedure: The forward
committor $q^\fwd_j(i)$ is defined as the probability of visiting coreset
$C_j$ next conditioned on being at state $i$.  The
backward committor $q^\bwd_j(i)$  is defined as the probability
of last coming from $C_j$ conditioned on being at state $i$.

% \subsection{Building the Markov State Model from the NEMD simulation}
\noindent\textbf{Remark 2:}
Sometimes, it is not easy to have an explicit expression for the committors, however, we
can still build the MSM by using the following procedure: Firstly compute the
matrixes $\hat T$ and $\hat M$ by
\begin{align}\label{eq:msm-ht}
  \hat M_{jk} &= P [\,h_{mT}(C_k) < h_{mT}(\cup_{l\neq k}C_l) \,\vert\, \hat x_{mT} = j\,],\\\label{eq:msm-hm}
  \hat T_{jk} &= P [\,h_{mT+T}(C_k) < h_{mT+T}(\cup_{l\neq k}C_l) \,\vert\, \hat x_{mT} = j\,],
\end{align}
where $\hat x_{mT}$ is milestone process corresponding to the NEMD trajectory $x_{mT}$ (notice here
only integral periods $mT$ along the NEMD trajectories are used for building the milestone process).
In Eq.~\eqref{eq:msm-ht} and \eqref{eq:msm-hm}, the time parameterized first hitting time of  set $A$ is defined by:
\begin{align}
  h_{t}(A) = \inf_{s\geq t}\{ s : x_s \in A \}.
\end{align}
It can be shown that (see Appendix~\ref{sec:app-prove} for details):
\begin{align}\label{eq:msm-ht-1}
  \hat T_{jk} &= \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)},\\\label{eq:msm-hm-1}
  \hat M_{jk} &= \frac{\langle q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)}.
\end{align}
% \recheck{We may need to write why do these formula works, because in literature it is only available for reversible processes.}
Therefore we have
\begin{align}\label{eq:msm-tmp26}
  \hat T_{ij} - \hat M_{ij}
  = \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  - \frac{\langle q_j^\bwd,  q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \frac{\langle (P^b - \id) q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu(j)}
  = \hat P_{jk}, \quad j\neq k
\end{align}
By Eq.~\eqref{eq:msm-ht} and \eqref{eq:msm-hm}, we have $\sum_i\hat M_{ij} = \sum_i\hat T_{ij} = 1$, then
$\sum_k(\hat M_{jk}- \hat T_{jk}) = 0$. By the second line of \eqref{eq:msm-tmatrix-00}, $\sum_k\hat P_{jk} = 1$, so using \eqref{eq:msm-tmp26} we have
$\hat T_{jj} - \hat M_{jj} + 1 = \hat P_{jj}$.
In the matrix expression the MSM transition matrix $\hat P$ is given by
\begin{align}\label{eq:msm-tmatrix-10}
  \hat P = \hat T - \hat M + \id
\end{align}
% Given coresets $\{C_1, \cdots, C_k\}$, the procedure \eqref{eq:msm-ht}--\eqref{eq:msm-tmatrix-10}
% If assuming that we have the coresets $\{C_1, \cdots, C_k\}$, no matter how they are identified,
In constrast to Eq.~\eqref{eq:msm-tmatrix-00}, which needs the intermediate spacial and
temporal discretization (i.e.~the Floquet transition matrix $P$), 
Eq.~\eqref{eq:msm-ht}, \eqref{eq:msm-hm} and \eqref{eq:msm-tmatrix-10} allows a direct derivation of the MSM from the original NEMD
trajectories, which substantially simplified the MSM building.
However, in the present paper, we present both the approachs,
because the former contains more information regarding the original dynamics in $P$ and $L(t)$ than in $\hat P$.
Moreover, the coreset identification may need a good spacial discretization to the original NEMD trajectories, which is
usually consisten with that used to go from $\ml(T)$ to $L(t)$. In the end, the MSM
built from both approachs should be consistent, taking no account of the statistical uncertainty in the samplings.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Proves for Equations~\eqref{eq:msm-ht-1} and \eqref{eq:msm-hm-1} }
\label{sec:app-prove}
